---
title: "GBIF Preprocessing"
author: "J. Kluge, H. Meyer, T. Nauss"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GBIF Preprocessing}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo=FALSE}
library(data.table)

inpath <- "D:/active/juergen/"
datapath <- paste0(inpath, "data/")
datapath_gbif <- paste0(datapath, "gbif/")
datapath_rdata <- paste0(datapath, "rdata/")

load(paste0(datapath_rdata, "gibf_01_initial_input.Rdata"))
```


## Reading GBIF data dump for the first time
Reading GBIF for the first time (code is not evalueted, but the saved file
will be loaded in the end).
```{r, eval=FALSE}
# Read GBIF for the first time -------------------------------------------------
# Split GBIF into chunks and load them into a data.table afterwards.
# Splitting is done because there was an error once regarding the correct column
# numbers. Only columns given by relevant_cols will be processed during the 
# final import.
infile <- paste0(datapath_gbif, "gbif_PTERIDOPHYTA.txt")
  splitLTF(infile, sep = "\t")
  
  infiles <- list.files(datapath_gbif, pattern = glob2rx("gbif_chunk*.txt"),
                        full.names = TRUE)
  relevant_cols <- c(1, 63, 70, 71, 72, 78, 79, 93, 100, 157, 164, 173, 182, 
                     210, 213, 214, 215, 216, 217, 218, 219, 220)
  gbif <- readLTF(infiles, sep = "\t", rlvt_cols = relevant_cols)  
  
  for(i in seq(40)){
    print(dim(gbif[[i]]))
  }
  
  gbif <- gbif[-1,]
  save(gbif, file = paste0(datapath_rdata, "gibf_01_initial_input.Rdata"))
```
As it turns out, everything worked fine and the `gibf_01_initial_input.Rdata` 
dataset will be used from now on.


## Clean GBIF data dump
First check how many cells have been empty in the original data set.
```{r}
cell_empty <- sapply(names(gbif), function(x){
  gbif[, length(which(gbif[[x]] == "cempty"))]
  })
cell_empty
```

Let's check, how many cells have been NA values (or interpreted as such.
```{r}
cell_na <- sapply(names(gbif), function(x){
  gbif[, length(which(is.na(gbif[[x]])))]
  })
cell_na
```

Time for a closer look on these NA's in column `countryCode`.
```{r}
gbif[which(is.na(countryCode)), 1:7, with = FALSE]
```
In total, 1007 rows have already NA values. We come back to that later but at
this point we can be sure that e.g. `gbifID` 318859294 will not be interpreted
as e.g. Namibia for any country code conversion we will apply later.

Let's convert "cempty" (i.e the initially empty cells) to NA.
```{r}
for (i in seq_len(ncol(gbif))){
  set(gbif, i = which(gbif[[i]]== "cempty"), j = i, value = NA)
  }
cell_na <- sapply(names(gbif), function(x){
  gbif[, length(which(is.na(gbif[[x]])))]
  })
cell_na
```
Worked. Only column `countryCode` has a differing value between the cells which
used to be "cempty" and the ones which are now NA but the difference matches
the 1007 rows which had alread NA's in this column.

Let's finish the initial cleaning by converting the columns to their individual
class. At the moment, all columns are of class `character`.
```{r}
head(gbif)
```

Columns `gbifID`, `decimalLatitude` and `decimalLongitude` can be converted to 
numeric:
```{r}
for(i in c("gbifID", "decimalLatitude", "decimalLongitude")){
  set(gbif, j = i, value = as.numeric(gbif[[i]]))
  }
summary(gbif[, c("gbifID", "decimalLatitude", "decimalLongitude"), 
             with = FALSE])
```


## Create "geographical" subset of GBIF data dump
For the upcoming analysis, only those data lines are relevant which have at 
least either latitude and longitude information or a country code or county
information. Let's get an overview of that stuff:
```{r}
total <- nrow(gbif)

either <- gbif[, length(which(!is.na(countryCode) | !is.na(decimalLatitude) |
                              !is.na(county)))]

ctry_only <- 
  gbif[, length(which(!is.na(countryCode) & is.na(decimalLatitude) & 
                        is.na(county)))]
  
coord_only <- 
  gbif[, length(which(is.na(countryCode) & !is.na(decimalLatitude) & 
                        is.na(county)))]

cnty_only <- 
  gbif[, length(which(is.na(countryCode) & is.na(decimalLatitude) & 
                        !is.na(county)))]

ctry_and_coord <- 
  gbif[, length(which(!is.na(countryCode) & !is.na(decimalLatitude)))]

cnty_and_coord <- 
  gbif[, length(which(is.na(countryCode) & !is.na(decimalLatitude) & 
                        !is.na(county)))]

ctry_and_cnty <- 
  gbif[, length(which(!is.na(countryCode) & is.na(decimalLatitude) & 
                        !is.na(county)))]

no_geoinfo <- 
  gbif[, length(which(is.na(countryCode) & is.na(decimalLatitude) &
                        is.na(county)))]
```
```{r, echo=FALSE}
paste0("Total number of data lines:               ", total)
paste0("Country and/or county and/or coordinates: ", either)
paste0("Country only:                             ", ctry_only)
paste0("Coordinates only:                         ", coord_only)
paste0("County only:                              ", cnty_only)
paste0("Country and coordinates:                  ", ctry_and_coord)
paste0("Only county and coordinates:              ", cnty_and_coord)
paste0("Only country and county:                  ", ctry_and_cnty)
paste0("No information:                           ", no_geoinfo)
```
These figures are not bad. Overall, at least one minimum geographic information
is available for `r either` of the `r total` lines available in the GBIF data
dump which is equal to `r round(either/total,2)*100`%.

For `r gbif[, length(which(!is.na(decimalLatitude)))]` 
(`r round(gbif[, length(which(!is.na(decimalLatitude)))]/total,2)*100`%), 
geographical coordinates are availabe. Almost all of these observations 
(i.e. `r ctry_and_coord + cnty_and_coord` which equals all except `r coord_only`)
have an additional information on either the country or county, so that 
cross-validation between these two kinds of information is possible.

This reminds us with a total of `r ctry_only+cnty_only+ctry_and_cnty` (
`r round((ctry_only+cnty_only+ctry_and_cnty)/total,2)`%) which can only be 
assigned on a county or country level (aside from the 
`r round(no_geoinfo/total,2)*100`% which can not be geocoded at all). 

To finally get a first cleanded version of the GBIF data dump, the lines with
absolutely no information will be removed from the dataset (i.e. the rest will
be kept).
```{r}
gbif <- gbif[which(!is.na(countryCode) | !is.na(decimalLatitude) |
                       !is.na(county))]
```

All done. Let's store the current state of the dataset so we can come back if 
something happens in the next section.
```{r, eval=FALSE}
save(gbif, file = paste0(datapath_rdata, "gibf_02_cleaned_input.Rdata"))
```
