---
title: "GBIF Geocodation"
author: "J. Kluge, H. Meyer, T. Nauss"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GBIF Geolocation}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo=FALSE}
library(data.table)

inpath <- "D:/active/juergen/"
datapath <- paste0(inpath, "data/")
datapath_gbif <- paste0(datapath, "gbif/")
datapath_rdata <- paste0(datapath, "rdata/")
datapath_geo <- paste0(datapath, "geo/")

load(paste0(datapath_rdata, "gibf_02_cleaned_input.Rdata"))
ctry_and_coord <- 
  gbif[, length(which(!is.na(countryCode) & !is.na(decimalLatitude)))]
cnty_and_coord <- 
  gbif[, length(which(is.na(countryCode) & !is.na(decimalLatitude) & 
                        !is.na(county)))]
```


# Check GBIF coordinate information for plausibility
Since actual geolocation is quite relevant for the upcoming analysis, a cross
check of geographical coordinates and country/county information will be 
performed for the 'r ctry_and_coord + cnty_and_coord' datasets which have both 
information.
```{r}
cntr <- rgdal::readOGR(paste0(datapath_geo, "world_boundaries.shp"), 
                       layer = "world_boundaries")

```




## Geocode GBIF by country code information
After checking the coordinates for plausibility, missing coordinates will be
assigned by country code or county name. The data source for country information
is <http://www.gadm.org/version2>.

First, get the country centroids from the country data layer of GDAM2:
```{r}
test <- rgdal::readOGR(paste0(datapath_geo, "test.shp"), 
                       layer = "test")




cntr <- rgdal::readOGR(paste0(datapath_geo, "world_boundaries.shp"), 
                       layer = "world_boundaries")
cntr_cent <- calcPolygonCentroids(cntr)
```

Second, get the county centroids from the country data layer of GDAM2:
```{r}
county <- rgdal::readOGR(paste0(datapath_geo, "world_boundaries.shp"), 
                       layer = "world_boundaries")
county_cent <- calcPolygonCentroids(county)
```


## Vignette Info

Note the various macros within the `vignette` setion of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
